---
title: "Data Preprocessing"
author: "Iker González Cuellar"
date: "07-11-22"
output:
  html_document: default
---

```{r}

library(dplyr)

library(magrittr)

setwd('D:/iCloud/iCloudDrive/Master Big Data/Data Analytics - Teoria_Practica/_r_exercises/3-DataPreprocessing')

BcnPisos = read.delim("_repository/bcn_pisos.txt")
```

#### 1. Let's play a bit with pipes. Using the pipeline operator perform the following operations:

##### a) Compute the squared root of the squared of any number.

```{r}

x = 32 %>% .^2 %>% sqrt
print(x)
```

##### b) Sample 1000 individuals from a normal distribution (mean = 5 , sd = 3 ), standardize the sample (subtract the mean and divide by the standard deviation, i.e., scale) and compute the max value.

```{r}
y =  data.frame(rnorm(n= 1000, mean = 5, sd = 3))
colnames(y)[1] = "observations"

y["standardize_observations"] = (y$observations - mean(y$observations))/ sd(y$observations)

cat("max:",max(y$standardize_observations))

```

##### c) Same as b) but plotting

```{r}

x = rnorm(n= 1000, mean = 5, sd = 3)
hist(x, probability = T, breaks = 50, col = '#68f78e')
curve(dnorm(x,5,3), col = 'red', add = T)

```

#### 2. With the pisos dataset and using an only pipeline, compute the following transformations:

##### a) Drop the duplicated individuals and compute the mean value of the fats ("Valor") by district ("Dist").

```{r}
BcnPisosMeanbyDist = BcnPisos %>% extract() %>% unique(.) %$% aggregate(Valor ~ Dist, FUN = mean )

colnames(BcnPisosMeanbyDist)[2] = "Valor_medio"
```

```{r}
print(BcnPisosMeanbyDist)
```

##### b) Drop the duplicated individuals, get the numeric features of the dataset and standardize it.

```{r}
BcnPisosScaled = BcnPisos %>% extract(,1:5,9) %>% unique(.) %>% scale(.,center = T, scale = F)

head(BcnPisosScaled)
```

##### c) Drop the duplicated individuals, add a new factor to the dataset "Greater than is mean"with values (Y,N) indicating if the Value ("Valor") of the at is greater or not than the mean of the flats in the district.

```{r}

BcnPisosUniqueMeans = BcnPisos %>% unique(.) %>% left_join(., BcnPisosMeanbyDist, by = 'Dist') %>% mutate("Greater than is mean" = if_else(Valor > Valor_medio, "Yes", "No"))

head(BcnPisosUniqueMeans)
```

#### 3. Finally, you are asked to do a complete transformation of the pisos dataset.

#### We want to analyse and visualize some general features of the districts of the city, characterizing a sample of flats.

#### To ease this analysis you are asked to create a new dataset containing one individual for each district and containing the following variables:

##### a) Propose R code for the transformation of this dataset. You are free to use any technique explained during the course (and others) but the use of some pipes will be valued positively (7 points).

```{r}
library(reshape2)
library(tidyr)

# CODIFICACIÓN BIG IF PARA ETIQUETAR LAS VARIABLES EN SUB_VARIABLES DE ACUERDO AL LISTADO SOLICITADO

bigIf = function (x) {
  if_else(x$variable == "Valor", "Valor", 
          if_else(x$variable == "Superf", "Superf", 
                  if_else((x$variable == "Dorm" & x$value == 1), "1 Dorm", 
                          if_else((x$variable == "Dorm" & x$value == 2), "2 Dorm",
                                  if_else((x$variable == "Dorm" & x$value == 3), "3 Dorm",
                                          if_else((x$variable == "Dorm" & x$value == 4), "4 Dorm",
                                                  if_else((x$variable == "Dorm" & x$value == 5), "5 Dorm",
                                                          if_else((x$variable == "Ascens" & x$value == "SI"), "AscS",
                                                                  if_else((x$variable == "Ascens" & x$value == "NO"), "AscN",
                                                                          if_else((x$variable == "Planta" & x$value == "Atic"), "Atics",
                                                                                  if_else((x$variable == "Planta" & x$value == "Planta"), "Plantas",
                                                                                          if_else((x$variable == "Planta" & x$value == "Bajos"), "Bajos",
                                                                                                  if_else((x$variable == "Edat" & as.integer(x$value) <= 10), "Nous",
                                                                                                          if_else((x$variable == "Edat" & as.integer(x$value) >= 11 & as.integer(x$value) <= 20), "SemiNous",
                                                                                                                  if_else((x$variable == "Edat" & as.integer(x$value) >= 21 & as.integer(x$value) <= 50), "Vells",
                                                                                                                          if_else((x$variable == "Edat" & as.integer(x$value) >= 51), "MoltVells",
                                                                                                                                  "0 Unused Variables"
                                                                          ))))))))))))))))

}

# TRANSFORMACIÓN PASO 1:

BcnPisosTransform = BcnPisos[,c(1:3,5,7:8,11)] %>% unique(.) %>% melt(., id.vars = "Dist") %>% mutate("sub_variable" = suppressWarnings(bigIf(.)))             # se ignoran los warning arrojados por la función del BIG IF, puesto que se ha validado previamente la tabla y no se ha introducido ningún NA en el data frame trás su aplicación.

# TRANSFORMACIÓN PASO 2:

BcnPisosToMean = BcnPisosTransform[BcnPisosTransform$sub_variable == c("Valor","Superf"),] %>% .[,c(1,3,4)] %>% transform(., value = as.numeric(value)) %>% pivot_wider(names_from = sub_variable, values_from = value, values_fn = mean)


BcnPisosToCount = BcnPisosTransform[!(BcnPisosTransform$sub_variable %in% c("Valor","Superf")),] %>% .[,c(1,4)] %>% mutate("countObs" = 1) %>% pivot_wider(names_from = sub_variable, values_from = countObs, values_fn = sum, values_fill = 0, names_sort = T)

# TRANSFORMACIÓN PASO 3:

P1_BcnPisoSummary = full_join(BcnPisosToMean, BcnPisosToCount, by = "Dist")

print(P1_BcnPisoSummary)
```
