---
title: 'Time Series Analysis'
subtitle: ' 4.  Time Series for Count Data'
author: "Pedro Delicado. EIO, UPC"
date: "February 8th, 2023"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message=FALSE)
path.data.0 <- "../_repository/"
path.data.1 <- "../_repository/Datasets/" #To be replaced by "Time_Series_Datasets/" or the appropriate path directory
```

# Introduction and examples

------------------------------------------------------------------------

## Introduction to Time Series of Count Data

-   **Count variable**: It is a (random) variable that takes values at
    $\mathbb{N}_0=\{0,1,2,\ldots.\}$, or in a subset of this set.
-   Other types of **Discrete-Valued Time Series**:
    -   Ordinal time series.
    -   Nominal time series.
-   A good reference book:
    -   Weiss, C.H. (2018) *An Introduction to Discrete-Valued Time
        Series*. John Wiley & Sons, Inc., Chichester.
    -   [Companion website: datasets and source
        code](https://www.wiley.com/go/weiss/discrete-valuedtimeseries).
        (Use `DiscrValTS17`)
-   [R package `tscount`: An R Package for Analysis of Count Time Series
    Following Generalized Linear
    Models](https://CRAN.R-project.org/package=tscount).

------------------------------------------------------------------------

## Examples of count data time series

### Traffic accidents in Spain, 2014-2017

Data source: DGT, Direccion General de Trafico

-   Traffic accidents micro-data file.
-   Each line in the file corresponds to an accident with victims that
    occurred on interurban roads or in crossings.
-   Included variables: province, road and kilometric point in which the
    accident occurred and the resulting seriousness in those involved
    are included.
-   The data does not include accidents that occurred in Catalonia and
    the Basque Country.
-   The number of records is 104.542.
-   Download date: 22/04/2019.

------------------------------------------------------------------------

```{r,warning=FALSE,message=FALSE}
path.data <- paste0(path.data.0,"accidentes_trafico_2014_2017/")
library(readxl)
accidts <- read_xlsx(paste0(path.data,"ACC_CARRETERA_PK_2014_2017_F.xlsx"),
                     sheet="ACC_CARRETERA_PK_2014_2017_F")
# names(accidts)
#  [1] "NUM_ACC"            "ANYO"               "FECHA_ACCIDENTE"   
#  [4] "HORA_2F"            "PROVINCIA_1F"       "ZONA_2F"           
#  [7] "CARRETERA"          "PK"                 "SENTIDO_1F"        
# [10] "TITULARIDAD_VIA_2F" "TIPO_VIA_2F"        "TRAZADO_1F"        
# [13] "TIPO_ACCIDENTE_3F"  "TOTAL_MU30DF"       "TOTAL_HG30DF"      
# [16] "TOTAL_HL30DF"      

library(chron) # para usar las funciones month.day.year y day.of.week
aux <- month.day.year(as.Date(as.character(accidts$FECHA_ACCIDENTE),"%Y-%m-%d"))

accidts$Dia <- aux$day
accidts$Mes <- aux$month
accidts$Anyo <- aux$year
accidts$Dia.semana <- day.of.week(aux$month, aux$day, aux$year) # Sunday is "0"
accidts$Dia.semana[accidts$Dia.semana==0] <- 7 # Reemplazamos "0" por "7"
accidts$Fin.semana <- ifelse( (accidts$Dia.semana>5), 1, 0)

accidts$td <- julian(aux$month, aux$day, aux$year, 
                    origin.=c(month = 12, day = 31, year = 2013))

td.max <-  julian(x=12, d=31, y=2017, 
                  origin.=c(month = 12, day = 31, year = 2013))
# [1] 1461

dias <- 1:td.max # variable tiempo en dias, desde el 1-1-2014 hasta el 31-12-2017 
Year.continuous <- 2014+dias/365 # variable tiempo en anyos, con precision de dias

accidts$tm <- accidts$Anyo+(accidts$Mes-1)/12
meses <- seq(2014, 2017+11/12, by=1/12) # variable tiempo en meses

df2ts <- function(t, var.t, var=var.t, agg.FUN="length", 
                  default.value=NA, frequency=1){
# t: time index 
# var.t: variable with the time (in the scale of t)
# var: variable to be aggregated 
# agg: aggregate function
# default.value: the value in the time series for instants with no corresponding rows in the data frame
# frequancy:   
  len.t <- length(t)
  time.series <-rep(default.value,len.t)
  aux <- aggregate(var,by=list(var.t),FUN=agg.FUN)
  Id <- match(aux[[1]],t)
  time.series[Id] <- aux[[2]]
  time.series <- ts(data=time.series, start=t[1],end=t[len.t],frequency=frequency)
  return(time.series)
}
```

```{r,warning=FALSE,message=FALSE}
# Agregar por dias el numero de accidentes, heridos, heridos graves y leves:
acc.d <- ts(df2ts(dias, var.t=accidts$td),start=2014,frequency = 365)
mu.d <- ts(df2ts(dias, var.t=accidts$td, var=accidts$TOTAL_MU30DF,agg.FUN=sum),start=2014,frequency = 365)
hg.d <- ts(df2ts(dias, var.t=accidts$td, var=accidts$TOTAL_HG30DF,agg.FUN=sum),start=2014,frequency = 365)
hl.d <- ts(df2ts(dias, var.t=accidts$td, var=accidts$TOTAL_HL30DF,agg.FUN=sum),start=2014,frequency = 365)

plot(ts.union(acc.d,mu.d,hg.d,hl.d), main="Daily accidents, deaths, serious and minor injuries")
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.d, main="Daily deaths by traffic accidents")
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.d, main="Daily deaths by traffic accidents",type="b",pch=19)
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.d, main="Daily deaths by traffic accidents", type="p",pch=19)
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r}
# Agregar por meses el numero de accidentes, heridos, heridos graves y leves:
acc.m <- df2ts(meses, var.t=accidts$tm, frequency = 12)
mu.m <- df2ts(meses, var.t=accidts$tm, var=accidts$TOTAL_MU30DF,agg.FUN=sum, frequency = 12)
hg.m <- df2ts(meses, var.t=accidts$tm, var=accidts$TOTAL_HG30DF,agg.FUN=sum, frequency = 12)
hl.m <- df2ts(meses, var.t=accidts$tm, var=accidts$TOTAL_HL30DF,agg.FUN=sum, frequency = 12)

plot(cbind(acc.m,mu.m,hg.m,hl.m), main="Monthly accidents, deaths, serious and minor injuries")
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.m, main="Monthly deaths by traffic accidents",type="b",pch=19)
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r}
# Agregar por dias para una provincia
# el numero de accidentes, heridos, heridos graves y leves:
prov <- "Albacete"
Ipr <- which(accidts$PROVINCIA_1F==prov)
acc.d <- ts(df2ts(dias, var.t=accidts$td[Ipr]),start=2014,frequency = 365)
mu.d <- ts(df2ts(dias, var.t=accidts$td[Ipr],
                 var=accidts$TOTAL_MU30DF[Ipr],agg.FUN=sum),start=2014,frequency = 365)
hg.d <- ts(df2ts(dias, var.t=accidts$td[Ipr],
                 var=accidts$TOTAL_HG30DF[Ipr],agg.FUN=sum),start=2014,frequency = 365)
hl.d <- ts(df2ts(dias, var.t=accidts$td[Ipr],
                 var=accidts$TOTAL_HL30DF[Ipr],agg.FUN=sum),start=2014,frequency = 365)

plot(ts.union(acc.d,mu.d,hg.d,hl.d), main="Daily accidents, deaths, serious and minor injuries in Albacete")
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.d, main="Daily deaths by traffic accidents in Albacete")
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.d, main="Daily deaths by traffic accidents in Albacete",type="b",pch=19)
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.d, main="Daily deaths by traffic accidents in Albacete", type="p",pch=19)
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

```{r}
# Agregar por meses para una provincia
# el numero de accidentes, heridos, heridos graves y leves:
prov <- "Albacete"
Ipr <- which(accidts$PROVINCIA_1F==prov)
acc.m.pr <- df2ts(meses, var.t=accidts$tm[Ipr], frequency = 12)
mu.m.pr <- df2ts(meses, var.t=accidts$tm[Ipr], var=accidts$TOTAL_MU30DF[Ipr],agg.FUN=sum, frequency = 12)
hg.m.pr <- df2ts(meses, var.t=accidts$tm[Ipr], var=accidts$TOTAL_HG30DF[Ipr],agg.FUN=sum, frequency = 12)
hl.m.pr <- df2ts(meses, var.t=accidts$tm[Ipr], var=accidts$TOTAL_HL30DF[Ipr],agg.FUN=sum, frequency = 12)

plot(cbind(acc.m.pr,mu.m.pr,hg.m.pr,hl.m.pr), main="Monthly accidents, deaths, serious and minor injuries in Albacete")
```

------------------------------------------------------------------------

```{r,fig.asp=.6,fig.width=10,fig.height=6}
plot(mu.m.pr, main="Monthly deaths by traffic accidents in Albacete",type="b",pch=19)
abline(h=0,col=8)
abline(v=2014:2018,col=2,lty=2)
```

------------------------------------------------------------------------

*Example from the book of Weiss (2018)*

**Download counts.** Daily number of downloads of a TEX editor for the
period Jun. 2006 to Feb. 2007

```{r,fig.asp=.6,fig.width=10,fig.height=6}
Downloads <- ts(read.table(paste0(path.data.1,"Downloads.txt")))
plot(Downloads,type="b",pch=19,main="Download counts",ylab=expression(X[t]))
```

------------------------------------------------------------------------

*Example from the book of Weiss (2018)*

**Price stability in the Euro area (**$n=17$ countries). Monthly number
of countries having inflation rate below 2%, Jan. 2000 - Aug. 2012.

```{r,fig.asp=.6,fig.width=10,fig.height=6}
PriceStability <- ts(read.table(paste0(path.data.1,"PriceStability.txt")),
                           start=c(2000,1),frequency=12)
plot(PriceStability,type="b",pch=19,main="Price stability counts in EA17",ylab=expression(X[t]))
abline(v=2007,col=2,lty=2)
abline(h=0,col=8)
```

------------------------------------------------------------------------

*Example from the R library* `tscount`

**Campylobacter Infections Time Series.** Number of cases in the north
of the province Quebec (Canada) in four week intervals from January 1990
to the end of October 2000. It has 13 observations per year and 140
observations in total. Campylobacterosis is an acute bacterial
infectious disease attacking the digestive system.

```{r,fig.asp=.6,fig.width=6,fig.height=4}
library(tscount)
plot(campy) # campy is a ts object
abline(h=0,col=8)
```

````{=html}
<!---
***
*Example from the R library* `tscount`

**E. coli Infections Time Series.**
Escherichia coli (or E. coli) is a bacterium of the genus Escherichia that is commonly found in the lower intestine of warm-blooded organisms.
Most E. coli strains are harmless, but some serotypes can cause serious food poisoning in their hosts, and are occasionally responsible for food contamination incidents that prompt product recalls.
The  most virulent strains cause serious illness or death in the elderly, the very young, or the immunocompromised.

*** 

**E. coli Infections Time Series.**
Weekly number of reported disease cases caused by Escherichia coli in the state of North Rhine-Westphalia (Germany) from January 2001 to May 2013, excluding cases of EHEC (enterohaemorragic strains) and HUS (hemolytic-uremic syndrome).

```{r,fig.asp=.6,fig.width=6,fig.height=4}
#library(tscount)
ecoli.ts <- ts(ecoli$cases, start = c(ecoli$year[1],ecoli$week[1]), frequency = 52)
plot(ecoli.ts)
abline(h=0,col=8)
```
-->
````

# Characteristics of count data time series

------------------------------------------------------------------------

## Characteristics of count data time series

-   Let $X_1,\ldots,X_T$ be a count data time series:
    $X_t\in\mathbb{N}_0$.
-   **Visual description:** The standard time series plot.
-   **Numerical description (if stationary):**
    -   *Marginal characteristics:* Sample mean, sample variance.
    -   *Serial dependence:* ACF, PACF.
-   **Models**. ARIMA models are not appropriate because the
    **multiplication problem**: for a count random variable $X$ and for
    $\alpha\in \mathbb{R}$, usually
    $\alpha \cdot X \not\in \mathbb{N}_0$. Therefore,\
    $$
    X_t = \alpha X_{t-1} +\varepsilon_t \not\in \mathbb{N}_0
    $$ even if $X_{t-1}$ and $\varepsilon_t$ are both in $\mathbb{N}_0$.

# The INAR(1) model for count data time series

------------------------------------------------------------------------

## The INAR(1) model for count data time series (I)

-   The *Integer valued AR(1)* (or INAR(1)) model is an adaptation for
    count data of the real-valued AR(1) model, $$
    X_t=\alpha X_{t-1} + \varepsilon_t, \text{ where }  X_{t-1},  \varepsilon_t \in \mathbb{R} \text{ for all } t,
    $$
-   Another way to state the standard AR(1) model: $$
    \mathbb{E}(X_t|X_{t-1})= \alpha X_{t-1} 
    $$ because it is assumend that $\varepsilon_t$ and $X_{t-1}$ are
    independent and that $\mathbb{E}(\varepsilon_t)=0$.
-   Additionally, if $\varepsilon_t$ is assumed to be
    $N(0,\sigma^2_{\varepsilon})$, then the AR(1) model is equivalent to
    say that $$
    (X_t|X_{t-1}) \sim  N(\alpha X_{t-1}, \sigma^2_{\varepsilon}).
    $$

------------------------------------------------------------------------

## The binomial thinning operation

-   Let $k\in \mathbb{N}_0$ and let $\alpha\in (0,1)$, then we define $$
    \alpha \circ k =\sum_{i=1}^k Z_i
    $$ where $Z_1,\ldots,Z_k$ are i.i.d. Bernoulli r.v. with
    $\Pr(Z_i=1)=\alpha$. If $k=0$ then $\alpha \circ k=0$. So
    $\alpha \circ k \sim \text{Bin}(k,\alpha)$.
-   Let $X$ be a r.v. with values in $\mathbb{N}_0$, let
    $\alpha\in (0,1)$, and let $Z_1,\ldots,Z_k,\ldots$ i.i.d. Bernoulli
    r.v. with $\Pr(Z_i=1)=\alpha$ independent from $X$. Then we say that
    $$
    \alpha \circ X =\sum_{i=1}^X Z_i \sim \text{Bin}(X,\alpha)
    $$ is obtained from $X$ by **binomial thinning**.
-   **Interpretation:** Number of survivors from a population of size
    $X$, if individuals die independently of each other, with
    probability $1-\alpha$.

------------------------------------------------------------------------

## Binomial thinning

-   Observe that $\alpha \circ X\in \mathbb{N}_0$ and that $$
    \mathbb{E}(\alpha \circ X)=\alpha \mathbb{E}(X)
    $$ by the Law of Iterated Expectation: $$\mathbb{E}(\alpha \circ X)
    =\mathbb{E}(\mathbb{E}(\alpha \circ X|X))
    =\mathbb{E}(\alpha X)
    =\alpha \mathbb{E}(X).
    $$
-   Therefore the binomial thinning operation $\alpha\circ X$ is the
    integer-valued counterpart of the multiplication $\alpha X$.

------------------------------------------------------------------------

## The INAR(1) model

-   Let $\{\varepsilon_t\}_{t\in\mathbb{Z}}$ be a sequence of i.i.d.
    random variables with values in $\mathbb{N}_0$.
-   Let $\mu_{\varepsilon}=\mathbb{E}(\varepsilon_t)$ and
    $\sigma^2_{\varepsilon}=\text{Var}(\varepsilon_t)$.
-   Let $\alpha\in (0,1)$.
-   A stochastic process $\{X_t\}_{t\in\mathbb{Z}}$ which verifies that
    $$
    X_t=\alpha \circ X_{t-1} + \varepsilon_t \text{ for all } t \in \mathbb{Z}
    $$ is said to be an **INAR(1) process** if
    -   all thinning operations are performed independently of each
        other and of $\{\varepsilon_t\}_{t\in\mathbb{Z}}$, and
    -   the thinning at each time $t$, as well as $\varepsilon_t$, are
        independent of $\{X_s\}_{s<t}$.
-   **Interpretation:** Survivor process with immigration,

$\underbrace{X_t}_{\textrm{Population at time }t}$
$=\underbrace{\alpha \circ X_{t-1}}_{\textrm{Survivors of time }t-1}$
$+ \underbrace{\varepsilon_t}_{\textrm{Immigration}}$.

------------------------------------------------------------------------

## Properties of an INAR(1) process $\{X_t\}_{t\in\mathbb{Z}}$

$$
X_t=\alpha \circ X_{t-1} + \varepsilon_t 
$$

-   Markov property: $(X_t|X_{t-1},X_{t-2},\ldots) \sim (X_t|X_{t-1})$.
-   Transition probabilities: $$
    p_{k|\ell}=\Pr(X_t=k|X_{t-1}=\ell)=\sum_{j=0}^{\min\{k,\ell\}}\alpha^j(1-\alpha)^{\ell-j}
    \Pr(\varepsilon_t=k-j).
    $$
-   Conditional linear AR(1):
    $\mathbb{E}(X_{t}|X_{t-1})=\alpha X_{t-1}+\mu_{\varepsilon}$.
-   $\text{Var}(X_{t}|X_{t-1})=\alpha (1-\alpha)X_{t-1}+\sigma^2_{\varepsilon}$.
-   For second order stationarity (expectation $\mu$, variance
    $\sigma^2$ and autocorrelations $\rho(k)$ constant in $t$):
    -   $\mu=\frac{\mu_{\varepsilon}}{1-\alpha}$,
    -   $\sigma^2=\frac{\sigma^2_{\varepsilon}+\alpha \mu_{\varepsilon}}{1-\alpha^2}$,
    -   $\rho(k)=\alpha^k$, $k\ge 1$.

------------------------------------------------------------------------

## The Poisson INAR(1) Model

-   Assume that $\varepsilon_t \sim \text{Poisson}(\lambda)$. Then
    $\mu_{\varepsilon}=\sigma^2_{\varepsilon}=\lambda$.
-   It can be proved that there exists a unique **stationary marginal
    distribution** for $X_t$, and that this distribiution is also
    Poisson: $$
    X_t \sim \text{Poisson}\left(\mu=\frac{\lambda}{1-\alpha}\right).
    $$
-   That is, if $X_{t} \sim \text{Poisson}(\mu)$ then
    $X_{t+k} \sim \text{Poisson}(\mu)$ for all $k\ge 1$.
-   This follows from two properties of the Poisson distribution:
    -   If $X_1$ and $X_2$ are independent Poisson, the $X_1+X_2$ is
        also Poisson.
    -   If $X$ is Poisson, the $\alpha \circ X$ is also Poisson.

------------------------------------------------------------------------

-   For the Poisson INAR(1), the **index of dispersion** is equal to
    $1$: $$
    I=\frac{\text{Var}(X_t)}{\mathbb{E}(X_t)}=1.
    $$
-   We say that $X_t$ is *equidispersed*.
-   When *overdispersion* is present ($I>1$) then the **Negative
    Binomial INAR(1)** model could be more suitable than the Poisson.
-   Nevertheless, for simplicity here we only introduce the Poisson
    INAR(1) model.

------------------------------------------------------------------------

## Estimation in the INAR(1) model

-   A time series $x_1,\ldots,x_T$ is assumed to have been observed.
    from $X_1,\ldots,X_T$
-   Marginal mean $\mu$: Sample mean $\bar{x}$.
-   Marginal variance $\sigma^2$: Sample variance
    $\hat{\gamma}_0=\frac{1}{T}\sum_{t=1}^T(x_t-\bar{x})^2$.
-   Autocovariance of order $k$:
    $\hat{\gamma}_k=\frac{1}{T-k}\sum_{t=k+1}^T(x_t-\bar{x})(x_{t-k}-\bar{x})$.
-   Autocorrelation of order $k$:
    $\hat{\rho}_k = \hat{\gamma}_k/\hat{\gamma}_0$.
-   Estimation of $\alpha$ by the method of moments:
    $\hat{\alpha}_{MM}=\hat{\rho}_1$.

------------------------------------------------------------------------

-   Estimation in the Poisson INAR(1) model
    -   Model parameters: $\lambda$, $\alpha$.
    -   Estimation of $\lambda$ by the method of moments:
        $\hat{\lambda}_{MM}=\bar{x}(1-\hat{\alpha}_{MM})$.
    -   Other methods of estimation are possible: conditional least
        squares, maximum likelihood, conditional maximum likelihood.

------------------------------------------------------------------------

## Identification of the INAR(1) model

-   **Serial dependence:** Is the time series really following an
    INAR(1)?
    -   The ACF is like that of an AR(1)= $\rho(k)=\alpha^k$.
    -   The PACF is like that of an AR(1)=
        $\rho_{\text{part}}(1)=\alpha$ and $\rho_{\text{part}}(k)=0$ for
        $k>1$.
-   **Marginal distribution:** Is it a Poisson INAR(1)?
    -   Estimated dispersion index: $\hat{I}=\hat{\gamma}_0/\bar{X}$
    -   Reject the Poisson INAR(1) model at $0.05\%$ significance level
        when $$
          1\not \in \left(\hat{I} + \frac{1}{T}\frac{1+\hat{\alpha}_{MM}}{1-\hat{\alpha}_{MM}}
        \mp 1.96\sqrt{\frac{2}{T}\frac{1+\hat{\alpha}_{MM}^2}{1-\hat{\alpha}_{MM}^2}}\right).
          $$

------------------------------------------------------------------------

## Probability integral transform (PIT)

-   A more general identification tool: The **probability integral
    transform (PIT)**.
    -   It is defined from the assumed marginal model, combined with the
        observed data.
    -   It allows to plot the *PIT histogram*: It is expected to look
        like that of a uniform distribution when the assumed marginal
        model is correct.
    -   Otherwise, U shaped or inverse-U shaped histograms are observed.
-   Define
    $f_{k|\ell}=\Pr(X_t\le k|X_{t-1}=\ell)=\sum_{j=0}^k p_{j|\ell}$,
    where $p_{j|\ell}=\Pr(X_t=k|X_{t-1}=\ell)$ is computed assuming that
    the proposed marginal model is correct.

------------------------------------------------------------------------

-   For $u\in[0,1]$ define $$
    F_{t}(u)=\left\{
    \begin{array}{ll}
    0 & \text{if } u\le f_{x_t-1|x_{t-1}},  \\ 
    \frac{u-f_{x_t|x_{t-1}}}{f_{x_t|x_{t-1}}-f_{x_t-1|x_{t-1}}} & 
    \text{if } f_{x_t-1|x_{t-1}} < u < f_{x_t|x_{t-1}}, \\
    1 & \text{if } f_{x_t|x_{t-1}} \le u.
    \end{array}
    \right. 
    $$
-   Then define the *mean PIT* as $$
    \overline{F}(u)=\frac{1}{T-1}\sum_{t=2}^T F_t(u) \text{ for } u\in[0,1].
    $$ If the proposed marginal model is correct, then $\overline{F}(u)$
    should be close to the cumulative distribution function (c.d.f.) of
    a uniform distribution.
-   To check whether $\overline{F}(u)$ is the c.d.f. of a uniform
    distribution a histogram with $H$ bins is plotted (a value of $H=10$
    is usually appropriate), where bin $h$ has the height
    $\overline{F}(h/H)-\overline{F}((h-1)/H)$, $h=1,\dots,H$.

------------------------------------------------------------------------

## Forecasting in the INAR(1) model

-   We have observed $x_1,\ldots,x_T$ and we want to predict $X_{T+h}$,
    for some $h\ge 1$.
-   *Point forecast:* To predict the value of $X_{T+h}$. Minimizing the
    PMSE, the best prediction is $$
    \mathbb{E}(X_{T+h}|X_T=x_T) = \alpha^h x_T + \mu(1-\alpha^h).
    $$
-   *Conditional distribution forecasting:* To predict the complete
    conditional distribution of $(X_{T+h}|X_T=x_T)$, that is, $$
    p_{k|\ell}^{(h)}=\Pr(X_{T+h}=k|X_T=\ell), \text{ for all }k , \ell \in \mathbb{N}_0.
    $$
-   *Condtional variance* in the Poisson case: $$
    \text{Var}(X_{T+h}|X_T=x_T) = \alpha^h (1-\alpha^h) x_T + \mu(1-\alpha^h).
    $$

# Example 1: Download counts.

------------------------------------------------------------------------

*Download counts: Example from the book of Weiss (2018)*

Daily number of downloads of a TEX editor for the period Jun. 2006 to
Feb. 2007.

```{r}
#Fitting INAR(1) models to a given time series

nmax <- 50 #upper limit later used for approximate computations

#Transition probabilities of Poisson INAR(1) model:
tppois <- function(k,l,lambda,alpha){
	tp <- 0
	for(j in c(0:min(k,l))){
		tp <- tp + dbinom(j,l,alpha)*dpois(k-j,lambda)
	}
	#Alternatively, compute tp directly as
	#tp <- convolve(dbinom(0:k,l,alpha), rev(dpois(0:k,lambda)), type="open")[k+1]
tp
}


#Log-likelihood of Poisson INAR(1) model:
llpinar1 <- function(par,data){
#par is vector (lambda,alpha)
T <- length(data)
value <- -log(dpois(data[1], par[1]/(1-par[2]))) #full likelihood, otherwise use 0 here

for(t in c(2:T)) {
	value <- value-log(tppois(data[t], data[t-1], par[1], par[2]))
}
value
}


#i.i.d. Poisson model (par[1] is mean parameter):
llpiid <- function(par,data){
T <- length(data)
value <- 0

for(t in c(1:T)) {
	value <- value-log(dpois(data[t], par[1]))
}
value
}
```

```{r,fig.asp=.6,fig.height=6}
data <- read.table(paste0(path.data.1,"Downloads.txt"))[[1]]
Tlen <- length(data)
#Tlen #267

maxval <- max(data)
#maxval #14

plot(data, type="b", pch=19, cex=0.5, xlab = "t", ylab = expression("Download counts  x "[t]), cex.axis=0.85,
     main=paste(Tlen,"data"))
```

------------------------------------------------------------------------

```{r,fig.asp=.5,fig.width=10}
rho1 <- acf(data, plot=FALSE)[[1]][2]
#rho1 #0.2447806
op <- par(mfrow=c(1,2))
acf(data, lag.max=10, lwd=2, ci.col=gray(0.5))
legend("topright",paste("hat(alpha)=hat{rho}_1=",round(rho1,4)))

pacf(data, lag.max=10, lwd=2, ci.col=gray(0.5))
#AR(1)-like dependence structure
par(op)
```

------------------------------------------------------------------------

Analysis of marginal distribution

```{r}
hist(data-0.5, main="Histrogram of data")
```

------------------------------------------------------------------------

Relative frequencies

```{r}
absfreq <- tabulate(data+1) #+1 to include 0
plot(0:maxval, absfreq/Tlen, type="h", xlab = "k", ylab = expression(paste("estimated P(X"[t],"=k)")), lwd=4, ylim=c(0,0.3))
abline(h=0,col=8)
```

------------------------------------------------------------------------

Observations' mean: $\hat{\mu}=\bar{x}$

```{r}
barX <- mean(data)
barX
#2.400749
```

Innovations' mean: $\hat{\mu}_{\varepsilon}=\hat{\mu}(1-\hat{\alpha})$

```{r}
mue <- barX*(1-rho1)
mue
#1.813092
```

Observations' variance: $\hat{\sigma}^2=\hat{\gamma}_0$

```{r}
sX <- var(data)
#sX
#7.534285

(Tlen-1)/Tlen*sX
#7.506067
```

Innovations' variance:
$\hat{\sigma}^2_{\varepsilon}=(1-\hat{\alpha}^2)\hat{\sigma^2}-\hat{\alpha}\hat{\mu}_{\varepsilon}$

```{r}
se <- (1-rho1^2)*sX-rho1*mue
se
#6.639039
```

------------------------------------------------------------------------

## Testing for equidispersion:

Observed Index of Dispersion (ID): $\hat{I}=\hat{\gamma}_0/\bar{X}$

```{r}
ID <- (Tlen-1)/Tlen*sX/barX
ID #3.126552
```

```{r}
meanID <- 1-1/Tlen*(1+rho1)/(1-rho1)
#meanID #0.9938268

sdID <- sqrt(2/Tlen*(1+rho1^2)/(1-rho1^2))
#sdID #0.09189935
```

95% confidence interval for the ID:

```{r}
ID + -1/Tlen*(1+rho1)/(1-rho1) + c(-1,1)*qnorm(0.975)*sdID
```

If 1 does not belong to this interval then we can reject the null
hypothesis of equidispersion.

```{r}
IDe <- ID*(1+rho1)-rho1
#IDe #3.647091

#Zero frequency:
# p0 <- length(data[data==0])/Tlen
p0 <- absfreq[1]/Tlen
#p0 #0.2771536

#exp(-barX)
#0.09065003
```

```{r}
#i.i.d. Poisson
#Here, ML estimate of mu identical to mean, i.e., to barX
estml <- suppressWarnings(optim(c(barX), llpiid, method="L-BFGS-B", lower=c(0.0001), upper=c(9999), control=list(ndeps=c(1e-4)), data=data, hessian=TRUE))

muestml <- estml$par[[1]]
ofiest <- estml$hessian #inverse covariance
neglmax <- estml$value
estcov <- solve(ofiest)

#Estimates:
#c(muestml)
#2.400749

#Estimated standard errors:
#c(sqrt(diag(estcov)))
#0.09482389

#AIC and BIC:
AIC <- 2*neglmax+2*1
BIC <- 2*neglmax+log(Tlen)*1
#c(neglmax, AIC, BIC)
#660.4577 1322.9153 1326.5026
```

------------------------------------------------------------------------

Maximum likelihood estimation of the Poisson INAR(1) model

```{r}
#Poisson INAR1
estml <- suppressWarnings(optim(c(mue,rho1), llpinar1, method="L-BFGS-B", lower=c(0.0001,0.0001), upper=c(9999,0.9999), control=list(ndeps=c(1e-4,1e-4)), data=data, hessian=TRUE))

lambdaestml <- estml$par[[1]]
alphaestml <- estml$par[[2]]
ofiest <- estml$hessian #inverse covariance
neglmax <- estml$value
estcov <- solve(ofiest)
```

Estimates: $\hat{\lambda}$, $\hat{\alpha}$

```{r}
c(lambdaestml,alphaestml)
#1.9909829 0.1742646
```

Observations' Poisson parameter:

```{r}
muestml <- lambdaestml/(1-alphaestml)
muestml #2.411163
```

````{=html}
<!---
Estimated probability of $X_t$ being equal to $0$:
```{r}
exp(-muestml) #0.08971088
```
--->
````

------------------------------------------------------------------------

Comparing the observed relative frequencies with the corresponding
probabilities of a Poisson distribution with parameter equal to the
estimated one.

```{r}
plot(0:maxval,absfreq/Tlen, type="h", xlab = "k", ylab = expression(paste("P(X"[t],")")), lwd=4, ylim=c(0,0.35))
points((0:maxval)+.25,dpois(0:maxval,lambda =muestml ), type="h", lwd=4, col=gray(0.5))
abline(h=0,col=8)
legend("topright",c("Observed frequencies","Estimated marginal distribution"), col=c(1,8),lwd=4,lty=1)
```

```{r}
#Estimated standard errors:
#c(sqrt(diag(estcov)))
#0.10996153 0.03269754

#AIC and BIC:
AIC <- 2*neglmax+2*2
BIC <- 2*neglmax+log(Tlen)*2
#c(neglmax, AIC, BIC)
#644.4243 1292.8485 1300.0230
```

------------------------------------------------------------------------

Pearson residuals: $$
e_t = \frac{x_t - \mathbb{E}(X_t|X_{t-1})}{\sqrt{\text{Var}(X_t|X_{t-1})}}, \, t=2,\ldots, T.
$$

```{r}
cMeanP <- function(l,lambda,alpha){
	alpha*l+lambda
}

cVarP <- function(l,lambda,alpha){
	alpha*(1-alpha)*l+lambda
}


res <- (data[2:Tlen]-cMeanP(data[1:(Tlen-1)], lambdaestml,alphaestml)) /
  sqrt(cVarP(data[1:(Tlen-1)], lambdaestml,alphaestml))
```

```{r,fig.asp=.6}
par(mar=c(3, 3, 0.5, 0.5), mgp=c(1.8, 0.6, 0))
par(mfrow=c(2,1))
plot(data, type="l", xlab = "t", ylab = expression("Download counts  x "[t]), cex.axis=0.85)
points(data, pch=19, cex=0.5)
lines(cMeanP(data[1:(Tlen-1)], lambdaestml,alphaestml),col=4,lwd=2)
legend("topleft",c("Raw data","Fitted values"),col=c(1,4),lwd=c(1,2),pch=c(19,NA),cex=.75)

plot(res, ylab = "Pearson residuals")
abline(h=0,col=8)
```

------------------------------------------------------------------------

ACF and PACF of the Pearson residuals

```{r}
op<-par(mfrow=c(1,2))
acf(res)
pacf(res)
#mean(res)
#-0.03693506
#var(res)
#2.871406
#min(res)
#-2.117143
#max(res)
#6.730956
par(op)
```

------------------------------------------------------------------------

Probability integral transform (PIT)

```{r}
#Matrix of all CDFs:
allcdfs <- array(0, c(maxval+2,maxval+1)) #additional zero row corresponding to F(-1)

for(l in c(0:maxval)){
	cpmf <- rep(0, (maxval+1))
	for(k in c(0:maxval)){
		cpmf[k+1] <- tppois(k,l,lambdaestml,alphaestml)
	}
	allcdfs[(2:(maxval+2)),l+1] <- cumsum(cpmf)
}

nobins <- 10
PIT <- array(0, c(2,nobins+1))

for(j in c(1:nobins)){
	u <- j/nobins
	pitval <- 0
	
	for(t in c(2:Tlen)){
		if(allcdfs[(data[t]+1), (data[t-1]+1)]<u){
			if(allcdfs[(data[t]+2), (data[t-1]+1)]<u){
				pitval <- pitval+1
			}else{
				pitval <- pitval+ (u-allcdfs[(data[t]+1), (data[t-1]+1)])/(allcdfs[(data[t]+2), (data[t-1]+1)]-allcdfs[(data[t]+1), (data[t-1]+1)])
			}
		}
	}
	PIT[1,j+1] <- pitval/(Tlen-1)
	PIT[2,j+1] <- PIT[1,j+1]-PIT[1,j]
}
#PIT
     # [,1]      [,2]      [,3]      [,4]       [,5]       [,6]       [,7]
# [1,]    0 0.2603143 0.3804455 0.4724656 0.55379076 0.61407046 0.67272706
# [2,]    0 0.2603143 0.1201312 0.0920201 0.08132514 0.06027969 0.05865661
           # [,8]       [,9]      [,10]     [,11]
# [1,] 0.72488214 0.77343326 0.83164187 1.0000000
# [2,] 0.05215508 0.04855111 0.05820862 0.1683581

PIT.freq= as.vector(rep(((1:nobins)-0.5)/nobins, PIT[2,2:(nobins+1)]*1000))
PIT.hist <- hist(PIT.freq, plot=FALSE, breaks=nobins)
PIT.hist$counts <- PIT.hist$counts/1000
plot(PIT.hist, freq=TRUE, main="PIT histogram", ylab="PIT histogram", xlab="u", col="gray")
```

------------------------------------------------------------------------

## Forecasting based on the fitted model:

Last observation: $X_T=$ `r data[Tlen]`

The h-step-ahead forecasting distribution is obtained from the
transition probabilities.

```{r,fig.asp=.6,fig.height=4}
hmax <- 5
forecasts <- array(0, c(hmax+1,maxval+1))
for(h in c(1:hmax)){
	for(k in c(0:maxval)){
		forecasts[h,k+1] <- tppois(k,data[Tlen],muestml*(1-alphaestml^h),alphaestml^h)
	}
	}
#Stationary marginal distribution:
forecasts[hmax+1,] <- dpois(c(0:maxval),muestml)

par(mfrow=c(1,2))

plot(0:maxval,forecasts[1,], type="h", xlab = "k", 
     ylab = expression(paste("P(X"[T+h],"=k | X"[T],"=7)")), 
     lwd=4, ylim=c(0,0.35))
points((0:maxval)+.25,forecasts[hmax+1,], type="h", lwd=4, col=gray(0.5))
abline(h=0,col=8)
legend("topright",c(paste("h =",1,"step ahead"), "Marginal distribution"), col=c(1,8),lwd=4,lty=1)

plot(0:maxval,forecasts[hmax,], type="h", xlab = "k", 
     ylab = expression(paste("P(X"[T+h],"=k | X"[T],"=7)")), 
     lwd=4, ylim=c(0,0.35))
points((0:maxval)+.25,forecasts[hmax+1,], type="h", lwd=4, col=gray(0.5))
abline(h=0,col=8)
legend("topright",c(paste("h =",hmax,"step ahead"), "Marginal distribution"), col=c(1,8),lwd=4,lty=1, cex=.75)
```

------------------------------------------------------------------------

### Point predictions

$\mathbb{E}(X_{T+h}|X_T=x_T) = \alpha^h x_T + \mu(1-\alpha^h)$,
$\text{Var}(X_{T+h}|X_T=x_T) = \alpha^h (1-\alpha^h) x_T + \mu(1-\alpha^h)$.

```{r}
hmax <- 5 
h <- 1:hmax
xT <- data[Tlen]
cond.mean.h <- alphaestml^h*xT + muestml*(1-alphaestml^h)
cond.var.h <- alphaestml^h*(1-alphaestml^h)*xT + muestml*(1-alphaestml^h)
#Stationary marginal mean: muestml
```

```{r,fig.asp=.6}
par(mar=c(3, 3, 0.5, 0.5), mgp=c(1.8, 0.6, 0))
plot(data, type="l", xlab = "t", ylab = expression("Download counts  x "[t]), 
     cex.axis=0.85, xlim=c(1,Tlen+hmax))
points(data, pch=19, cex=0.5)
abline(h=muestml,col=2)
abline(v=Tlen,col=3,lty=2)
fitted.values <-cMeanP(data[1:(Tlen-1)], lambdaestml,alphaestml)
lines(2:Tlen,fitted.values,col=4,lwd=2)

arrows(x0 = Tlen+h,
       y0 = cond.mean.h - 1.96* sqrt(cond.var.h),
       y1 = cond.mean.h + 1.96* sqrt(cond.var.h),
       angle = 90, code = 3, length = 0.04, col = "darkgrey", lwd = 2)

lines(Tlen+c(0,h), c(fitted.values[Tlen-1],cond.mean.h), col=6, lwd=2)

legend("topleft",
       c("Raw data","Fitted values","Predicted values","Estimated marginal mean",
         "Approx. 95% CI"),
       col=c(1,4,6,2,8),
       lwd=c(1,2,2,1,2),
       pch=c(19,NA,NA,NA,NA),
       cex=.75)
```

------------------------------------------------------------------------

```{r,fig.asp=.6}
par(mar=c(3, 3, 0.5, 0.5), mgp=c(1.8, 0.6, 0))
plot(data, type="l", xlab = "t", ylab = expression("Download counts  x "[t]), 
     cex.axis=0.85, xlim=c(180,Tlen+hmax))
points(data, pch=19, cex=0.5)
abline(h=muestml,col=2)
abline(v=Tlen,col=3,lty=2)
fitted.values <-cMeanP(data[1:(Tlen-1)], lambdaestml,alphaestml)
lines(2:Tlen,fitted.values,col=4,lwd=2)

arrows(x0 = Tlen+h,
       y0 = cond.mean.h - 1.96* sqrt(cond.var.h),
       y1 = cond.mean.h + 1.96* sqrt(cond.var.h),
       angle = 90, code = 3, length = 0.04, col = "darkgrey", lwd = 2)

lines(Tlen+c(0,h), c(fitted.values[Tlen-1],cond.mean.h), col=6, lwd=2)

legend("topleft",
       c("Raw data","Fitted values","Predicted values","Estimated marginal mean",
         "Approx. 95% CI"),
       col=c(1,4,6,2,8),
       lwd=c(1,2,2,1,2),
       pch=c(19,NA,NA,NA,NA),
       cex=.75)
```

# Count Time Series Following Generalized Linear Models

------------------------------------------------------------------------

-   The R package `tscount` provides likelihood-based estimation methods
    for analysis and modeling of *count time series* following
    *generalized linear models*.
-   Let $\{X_t:t\in\mathbb{Z}\}$ be a count time series.
-   Let $\{\boldsymbol{Z}_t:t\in\mathbb{N}\}$ be a time-varying
    $r$-dimensional covariate vector, say
    $\boldsymbol{Z}_t=(Z_{t,1},\dots,Z_{t,r})^\top$.
-   Let ${\cal F}_t$ be the history of the joint process
    $\{X_t,\boldsymbol{Z}_{t+1}:t\in\mathbb{Z}\}$ up to time $t$
    including the covariate information at time $t+1$.
-   For $t\in \mathbb{Z}$, let
    $\lambda_t=\mathbb{E}\left(X_{t}|{\cal F}_{t-1}\right)$ be the
    conditional mean of $X_t$ given the previous history.

------------------------------------------------------------------------

-   The general form of the model is $$
    g(\lambda_t) = \beta_0
      + \sum_{k=1}^p \beta_k \,\widetilde{g}(X_{t-k})
      + \sum_{\ell=1}^q \alpha_{\ell} g(\lambda_{t-{\ell}})
      + \boldsymbol{\eta}^\top \boldsymbol{Z}_t,
    $$ where $g:\mathbb{R}^+\rightarrow\mathbb{R}$ is a link function
    and $\widetilde{g}:\mathbb{N}_0\rightarrow\mathbb{R}$ is a
    transformation function.
-   The parameter vector $\boldsymbol{\eta}=(\eta_1,\dots,\eta_r)^\top$
    corresponds to the effects of covariates.
-   In the terminology of GLMs we call $\nu_t=g(\lambda_t)$ the linear
    predictor.
-   The distribution of $X_t$ given ${\cal F}_{t-1}$ is usually assumed
    to be Poisson or Negative Binomial.

------------------------------------------------------------------------

## The log-linear model of order $p$ and $q$ with covariates

-   Consider the general model with the logarithmic link function
    $g(x)=\log(x)$, $\widetilde{g}(x)=\log(x+1)$.
-   Using $\nu_t=\log(\lambda_t)$, we obtain that $$
    \nu_t = \beta_0
      + \sum_{k=1}^p \beta_k \,\log(X_{t-k}+1)
      + \sum_{\ell=1}^q \alpha_{\ell} \nu_{t-{\ell}}
      + \boldsymbol{\eta}^\top \boldsymbol{Z}_t.
    $$
-   Here we assume that the conditional distribution of $X_t$ given
    ${\cal F}_{t-1}$ is Poisson.

------------------------------------------------------------------------

### Estimation in the log-linear model

-   To ensure stationarity and ergodicity, the parameter space is taken
    to be $$
    \Theta = \Biggl\{
    \boldsymbol{\theta}\in\mathbb{R}^{p+q+r+1}:\ 
    |\beta_1|,\dots,|\beta_p|, |\alpha_1|,\dots,|\alpha_q|<1,
    \Biggr.
    $$ $$
    \Biggl.
    \left|\sum\limits_{k=1}^p \beta_k + \sum\limits_{\ell=1}^q \alpha_{\ell} \right| < 1
    \Biggr\}.
    $$
-   The estimation in `tscount` is done by *conditional maximum
    likelihood* estimation.\
-   In practice, the function `tscount::tsglm` implements the estimation
    process.
-   The output of this function is an object of class `tsglm`, for which
    there are several implemented methods:
    -   `print`, `summary`, `plot`,\
    -   `fitted`, `residuals`, `predict`,
    -   `coef`, `se`,
    -   `pit` (probability integral transform, PIT), among other.

# Example 2. Campylobacter Infections

------------------------------------------------------------------------

-   Campylobacterosis is an acute bacterial infectious disease attacking
    the digestive system.
-   Number of cases in the north of the province Quebec (Canada) in four
    week intervals from January 1990 to the end of October 2000.
-   It has 13 observations per year and 140 observations in total.

```{r,fig.width=6,fig.height=3}
library(tscount)
par(mar=c(3, 3, 0.5, 0.5), mgp=c(1.8, 0.6, 0))
plot(campy, ylab="Number of cases") # campy is a ts object
points(campy,pch=19,cex=.5)
abline(h=0,col=8)
abline(v=seq(1990,2001),col=2,lty=2)
```

------------------------------------------------------------------------

-   We fit a log-linear model to this time series using the function
    `tsglm`.
-   We select the logarithm link function, using the argument `link`.
-   For taking into account serial dependence we include a regression on
    the previous observation (an AR(1) term).
-   Seasonality is captured by regressing on $X_{t-13}$, the observed
    data 13 time units (which is about one year) back in time.
-   The aforementioned specification of the model for the linear
    predictor is assigned by the argument `model`.
-   We fit a Poisson conditional distribution, specifying it by the
    argument `distr`.

```{r, echo=TRUE}
campyfit_pois <- tsglm(campy, 
                       model=list(past_obs=c(1,13)), 
                       link="log", distr="poisson")
```

------------------------------------------------------------------------

```{r}
summary(campyfit_pois)
```

------------------------------------------------------------------------

Accordingly, the fitted model for the number of new infections $X_t$ in
time period $t$ is given by
$X_t|{\cal F}_{t-1} \sim \text{Poisson}(\lambda_t)$ with $$
\log(\lambda_t) = `r round(coef(campyfit_pois)[1],2)` + 
`r round(coef(campyfit_pois)[2],2)` X_{t-1} + 
`r round(coef(campyfit_pois)[3],2)` X_{t-13}, 
\quad t=1,\dots,`r campyfit_pois$n_obs`.
$$

------------------------------------------------------------------------

```{r}
par(mfrow=c(2,1))
par(mar=c(3, 3, 0.5, 0.5), mgp=c(1.8, 0.6, 0))
plot(campy, ylab="Number of cases") # campy is a ts object
points(campy,pch=19,cex=.5)
abline(h=0,col=8)
abline(v=seq(1990,2001),col=2,lty=2)
lines(fitted(campyfit_pois),col=4,lwd=2)
legend("topleft",c("Raw data","Fitted values"),col=c(1,4),lwd=c(1,2),pch=c(19,NA),cex=.75)

plot(residuals(campyfit_pois), ylab = "Response residuals")
abline(h=0,col=8)
```

------------------------------------------------------------------------

```{r}
par(mfrow=c(2,2))
acf(residuals(campyfit_pois), main ="")# "ACF of response residuals")
pacf(residuals(campyfit_pois), main ="")#  "PACF of response residuals")
pit(campyfit_pois, main = "PIT Poisson")
```

------------------------------------------------------------------------

**Predicting campylobacter infections**

-   We fit now the log-linear model until December 1998.
-   Then we predict the number of campylobacter infections from January
    1999 to October 2000.
-   We finally compare the model predictions with the true observed
    values.

```{r, echo=TRUE}
campy_until_1998 <- window(campy, end = c(1998,13))
campy_from_1999 <- window(campy, start = c(1999,1), 
                          end = end(campy))
campyfit_pois_98 <- tsglm(campy_until_1998, 
                          model=list(past_obs=c(1,13)), 
                          link="log", distr="poisson")
```

------------------------------------------------------------------------

```{r}
summary(campyfit_pois_98)
```

------------------------------------------------------------------------

```{r, echo=TRUE}
campy_pred_99<-predict(campyfit_pois_98, 
                  n.ahead = length(campy_from_1999), 
                  level = 0.9, global = FALSE, B = 2000)
campy_pred_99_global <- predict(campyfit_pois_98, 
                  n.ahead = length(campy_from_1999), 
                  level = 0.9, global = TRUE, B = 2000)
```

------------------------------------------------------------------------

```{r, fig.height=3, fig.width=7}
par(mar=c(3, 3, 0.5, 0.5), mgp=c(1.8, 0.6, 0))
plot(campy, ylab="Number of cases") # campy is a ts object
points(campy,pch=19,cex=.5)

lines(fitted(campyfit_pois_98), col=4, lty = "longdash", lwd=2)

arrows(x0 = time(campy_pred_99_global$interval),
       y0 = campy_pred_99_global$interval[, "lower"],
       y1 = campy_pred_99_global$interval[, "upper"],
       angle = 90, code = 3, length = 0.04, col = "pink", lwd = 2)

arrows(x0 = time(campy_pred_99$interval),
       y0 = campy_pred_99$interval[, "lower"],
       y1 = campy_pred_99$interval[, "upper"],
       angle = 90, code = 3, length = 0.04, col = "darkgrey", lwd = 2)

points(campy_from_1999, pch = 19) 
lines(x = c(time(campy_until_1998)[117], time(campy_pred_99$pred)), 
      c(fitted(campyfit_pois_98)[117], campy_pred_99$pred), 
      col = 4, lty = "solid", lwd = 3)

legend("topleft",
       c("Raw data", "Fitted values", "Predicted values", 
         "90% CI individual coverage", "90% CI global coverage"),
       col=c(1,4,4,"darkgrey","pink"),
       lwd=c(1,2,3,2,2),
       lty=c(1,2,1,1,1),
       pch=c(19,NA,NA,NA,NA),
       cex=.75)

```

------------------------------------------------------------------------

## References and links

-   Shumway, R.H. and D.S. Stoffer (2017) *Time Series Analysis and Its
    Applications. With R Examples* Fourth Edition. Springer. [Companion
    website: datasets and source
    code](https://www.wiley.com/go/weiss/discrete-valuedtimeseries).
-   [R package `tscount`: An R Package for Analysis of Count Time Series
    Following Generalized Linear
    Models](https://CRAN.R-project.org/package=tscount).
