---
title: "Data preprocessing"
author: "Miguel Albertí Pons"
date: "05-11-22"
output:
  html_document: default
---

```{r}
# import de las librerias necesarias 
library(dplyr)
```

------------------------------------------------------------------------

## Ejercicio 1

**Let's play a bit with pipes. Using the pipeline operator perform the following operations:**

Compute the squared root of the squared of any number.
------------------------------------------------------------------------

```{r}
number = 5
number^2 %>% sqrt() 
```
------------------------------------------------------------------------

Sample 1000 individuals from a normal distribution (mean = 5 ,sd = 3), 
standardize the sample (subtract the mean and divide by the standard 
deviation, i.e., scale) and compute the max value.
------------------------------------------------------------------------

```{r}

standard_sample_function = function(sample){
  (sample - mean(sample))/sd(sample)
  }
rnorm(n=10000,mean=5,sd=3) %>% standard_sample_function() %>% max()
          
```
------------------------------------------------------------------------
Same as b) but plotting the density function before computing the max value.
------------------------------------------------------------------------
```{r}
# Falta el max
rnorm(n=10000,mean=5,sd=3) %>% standard_sample_function() %>% 
  hist(., freq = FALSE)

```
------------------------------------------------------------------------

## Ejercicio 2

**With the pisos dataset and using an only pipeline, compute the follo-wing transformations:**

------------------------------------------------------------------------
```{r}

library(readr)
pisos_bcn = read.delim("bcn_pisos.txt")

```
------------------------------------------------------------------------
Drop the duplicated individuals and compute the mean value of the flats 
("Valor") by district ("Dist").
------------------------------------------------------------------------
```{r}
pisos_bcn %>% unique() %>% group_by(Dist) %>% 
  summarise(mean = mean(Valor), n = n())
```
------------------------------------------------------------------------
Drop the duplicated individuals, get the numeric features of the dataset and 
standardize it.
------------------------------------------------------------------------
```{r}
pisos_bcn %>% unique() %>% select_if(., is.numeric) %>%
  mutate_all(~(scale(.) %>% as.vector))
```
------------------------------------------------------------------------
Drop the duplicated individuals, add a new factor to the dataset "Greater than 
is mean"with values (Y,N) indicating if the Value ("Valor") of the flat is 
greater or not than the mean of the flats in the district.
------------------------------------------------------------------------

## Ejercicio 3

Finally, you are asked to do a complete transformation of the pisos
dataset. We want to analyse and visualize some general features of the
districts of the city, characterizing a sample of flats. To ease this analy-
sis you are asked to create a new dataset containing one individual for
each district and containing the following variables:
Propose R code for the transformation of this dataset. You are
free to use any technique explained during the course (and others)
but the use of some pipes will be valued positively.
------------------------------------------------------------------------

```{r}
get_new_data_frame = function(sample){
  new_data__frame = sample %>% group_by(Dist) %>% 
    summarise(mean_valor = mean(Valor),
              un_cinc_dorm = sum(Dorm >= 1 & Dorm <= 5),
              with_ascens = sum(Ascens == "SI"),
              without_ascens = sum(Ascens == "NO"),
              num_atic = sum(Planta == "Atic"),
              num_planta = sum(Planta == "Planta"),
              num_planta = sum(Planta == "Bajos"),
              num_nous = sum(Edat < 10),
              num_semi_nous = sum(Edat < 11 & Edat < 20),
              num_vells = sum(Edat > 21 & Edat < 50),
              num_molt_vells = sum(Edat > 51),
              mean_superf = mean(Superf),
              )
  return(new_data__frame)
}

pisos_bcn %>% get_new_data_frame()
```
------------------------------------------------------------------------

Propose nice visualizations of this new dataset (3 points).
------------------------------------------------------------------------
