---
title: "Linear Regression"
author: "Miguel Albertí Pons &Iker González Cuellar"
date: "09-01-23"
output:
  html_document:
    df_print: paged

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>
---

```{r}
library(dplyr)
library(reshape2)
library(magrittr)
```

## Ejercicio 1

**Read the data, create 4 univariate time series (IPC index, Monthly variation of IPC index, Yearly**

**variation of IPC index, Variation of IPC index from the begining of the year) and one 4-dimesional**

**time series as the union of the previous 4 (use ts.union). Plot the 4-dim time series.**

```{r}
ipcDataRaw = read.csv("_repository/INE_IPC.csv", sep = ",")

ipcFullTable = ipcDataRaw[6:19,]
ipcFullTable.ts = ts(ipcFullTable)

# IPC INDEX Transformación
----------------------------------
ipcIndex = ipcDataRaw[c(6,7),1:196]
names(ipcIndex) = ipcIndex[1,]
ipcIndex = ipcIndex[-1,] %>% rename(id = 1) %>% melt(., id = "id") %>% rename(date = 2, index = 3) %>% .[,2:3]
ipcIndex_2.ts = ts(ipcIndex[2]) 



plot(ipcIndex_2.ts, ylab="",main="IPC Index")
```

## Ejercicio 2

**Answer the following questions: i. Which series have a seasonal pattern? ii. Which series are stationary?**

```{r}

```

## Ejercicio 3

**Compute the lag 12 differences of the IPC time series (using function diff). Then plot the union of 3**

**time series: IPC, its lag 12 differences, and the series containing yearly variation of IPC index we read**

**from file INE_IPC.csv. Do you think that yearly variation of IPC index is stationary in variance?**

```{r}

```

## Ejercicio 4

**Compute the Inflation as Inflation \<- 100\*diff(IPC,lag=12)/lag(IPC,k=-12).**

**• Is Inflation of class ts?**

**• Why do you think Inflation is so similar to the yearly variation of IPC index?**

```{r}

```

## Ejercicio 5

**Compute and plot the Autocorrelation Functions (ACFs) of the following time series:**

**• The 4 we read from file INE_IPC.csv.**

**• The inflation time series.**

**• The IPC series after having taken consecutively differences with lags 1 (regular difference) and 12**

**(seasonal difference). Plot this series previously: diff( diff(IPC,lag=1), lag=12). Do you think it**

**is stationary?**

```{r}

```
